---
title: "Seurat Exercises"
output: html_notebook
---


```{r}
library(Seurat)
library(tidyverse)
library(SeuratData)

```

- We're going to be working with some data from human Pancreatic Islets. 
- Disclaimer, I have not run through this pipeline with this dataset before, so lets see how it goes
```{r}
data('panc8')
panc8
```

- to make the data more managable, we're going to subet it to about 25% of its original size
```{r}
cells <- colnames(panc8)
cells_ss <- sample(cells,3723)
panc8 <- panc8[,cells_ss]
```

- I recommend having the slides up as reference for the code 

## exercises

- calculate percent mtRNA for this data 

```{r}

```

- visualize it with an VlnPlot

```{r}

```

- keep cells that have less than 10% mt Expression

```{r}

```

- normalize the data 

```{r}

```

- find the 3000 most variable genes in the data 

```{r}

```

- plot variable features

```{r}

```

-  center and scale the data 

```{r}

```

- reduce dimensions with PCA

```{r}

```

- plot the PCA reduction

```{r}

```
- use ElbowPlot to idenity how many PC's to keep

```{r}

```

- cluster the data 

```{r}

```

- create a  UMAP reduction

```{r}

```

- plot UMAP reduction, coloring by clusters

```{r}

```

- calculate markers for each cluster

```{r}

```

- using VlnPlot, Ridgeplot, and FeaturePlot, explore the set of markers you generated

```{r}

```

