---
title: "Seurat Exercises"
output: html_notebook
---


```{r}
library(Seurat)
library(tidyverse)
library(SeuratData)
```

- We're going to be working with some data from human Pancreatic Islets. 
- Disclaimer, I have not run through this pipeline with this dataset before, so lets see how it goes

```{r}
data('panc8')

class(panc8@assays)
slotNames(panc8)

class(panc8@assays$RNA)

dim(panc8)
```

- to make the data more managable, we're going to subet it to about 25% of its original size
```{r}
cells <- colnames(panc8)
cells_ss <- sample(cells,3723)
panc8 <- panc8[,cells_ss]
```

- I recommend having the slides up as reference for the code 

## exercises

- calculate percent mtRNA for this data 

```{r}
panc8[["percent.mt"]] <- PercentageFeatureSet(panc8, pattern = "^MT-")
panc8[['ID']] <- 'ID'
panc8@meta.data
```

- visualize it with an VlnPlot

```{r}
VlnPlot(panc8, features = 'percent.mt', group.by = 'ID')
```

- keep cells that have less than 10% mt Expression

```{r}
panc8[['cell_id']] <- rownames(panc8@meta.data)
cells_that_pass_qc <- filter(panc8@meta.data, percent.mt < 10 ) %>% pull(cell_id)
panc8 <-  panc8[,cells_that_pass_qc]
```

- normalize the data 

```{r}
panc8  <- NormalizeData(panc8, normalization.method = "LogNormalize", scale.factor = 10000)

```

```{r}
panc8@assays
```

- find the 3000 most variable genes in the data 

```{r}
panc8 <- FindVariableFeatures(panc8, 
                             assay = 'RNA', 
                             selection.method = "vst", 
                             nfeatures = 3000)

panc8@assays$RNA@var.features
```

- plot variable features

```{r}
VariableFeaturePlot(panc8)

```

-  center and scale the data 

```{r}
panc8 <- ScaleData(panc8)
panc8@assays$RNA@scale.data
```

- reduce dimensions with PCA

```{r}
set.seed(101)
panc8 <- RunPCA(panc8, features = VariableFeatures(object = panc8))
```

- plot the PCA reduction

```{r}
DimPlot(panc8, reduction = 'pca', group.by = 'ID')
```
- use ElbowPlot to idenity how many PC's to keep

```{r}

ElbowPlot(panc8,ndims = 50)


```
- Looks like 12 is the number of PCs to keep

- cluster the data 

```{r}
panc8 <- FindNeighbors(panc8, dims = 1:12)
panc8 <- FindClusters(panc8, resolution = 0.4)
```

- create a  UMAP reduction

```{r}
panc8 <- RunUMAP(panc8, dims = 1:12)
```

- plot UMAP reduction, coloring by clusters

```{r}
DimPlot(panc8, reduction = "umap", group.by = 'seurat_clusters')
```

- calculate markers for each cluster

```{r}

panc8.markers <- FindAllMarkers(panc8, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

```

- using VlnPlot, Ridgeplot, and FeaturePlot, explore the set of markers you generated

```{r}

```

